git ---
title: "hotHandAnalysis"
author: "Ben Peloquin"
date: "May 6, 2016"
output: 
  html_document:
    toc: true
    toc_depth: 5
---

Prelims
```{r}
rm(list = ls())
library(dplyr)
library(tidyr)
library(ggplot2)
setwd('/Users/benpeloquin/Desktop/Spring2016/Stats267/hot_hand')
source("analysis_helpers.R")
path <- "player_data/"
```

# Section I - General data orientation

Read in data - see `get_data.py` for data collection code (first check store)
```{r}
## Note this takes appr 2min to run
d.raw <- read.csv("full_season_data_player_info.csv")

if (!inGlobalEnv("d.raw")) {
  files <- list.files(path)
  d.raw <- data.frame()
  for (f in files) {
    d <- read.csv(paste0(path, f))
    d.raw <- rbind(d.raw, d)
  }
  
  d.player_info <- read.csv("player_info/players_info.csv") %>%
    select(player_number, age, weight, height, years_in_league, player_id, position)
  names(d.player_info) <- c("PLAYER_NUMBER", "AGE", "WEIGHT", "HEIGHT", "YEARS_IN_LEAGUE", "PLAYER_ID", "POSITION")
  
  d.raw <- full_join(d.raw, d.player_info, by = 'PLAYER_ID')
  write.csv(d.raw, "full_season_data_player_info.csv")
}
```

# Data pre-processing

We have `r dim(d.raw)[1]` observations with `r dim(d.raw)[2]` predictor variables.

From the get go we can see a number of predictors we might want to recode. Looks like `GAME_ID`, `GAME_EVENT_ID`, `PLAYER_ID`, `TEAM_ID`, `PERIOD`, `SHOT_ATTEMPTED_FLAG`, and `SHOT_MADE_FLAG` should all be coded as factors (currently coded as ints). Let's also convert `SHOT_TYPE` to an integer value...
```{r}
d <- d.raw %>%
  mutate(GAME_ID = as.factor(GAME_ID),
         GAME_EVENT_ID = as.factor(GAME_EVENT_ID),
         PLAYER_ID = as.factor(PLAYER_ID),
         TEAM_ID = as.factor(TEAM_ID),
         PERIOD = as.factor(PERIOD),
         SHOT_VALUE = ifelse(SHOT_TYPE == "2PT Field Goal", 2, 3),
         TOTAL_TIME_REMAINING = SECONDS_REMAINING + 60 * MINUTES_REMAINING)
```

Let's do some preliminary exploring just to make sure this data makes sense...

# Basic data orientation / summary

## Total number of shots taken
```{r}
sum(player_shots_summary$total_shots_taken)
```

## Total number of shots made
```{r}
sum(player_shots_summary$total_shots_made)
```

## Number of players
```{r}
length(unique(d$PLAYER_ID)) == length(unique(d$PLAYER_NAME))
length(unique(d$PLAYER_ID))
```

## Number of teams
```{r}
length(unique(d$TEAM_ID))
```
Looks good, there are 30 teams

## Number of shots by player
```{r}
player_shots_summary <- d %>%
  group_by(PLAYER_NAME, TEAM_NAME) %>%
  summarize(total_shots_taken = n(),
            total_shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = total_shots_made / total_shots_taken)

ggplot(player_shots_summary, aes(x = total_shots_taken, y = total_shots_made, col = TEAM_NAME)) +
  geom_point(alpha = 0.5, size = 4) +
  geom_smooth(method = "lm", aes(group=1)) + 
  guides(col = FALSE) +
  ggtitle("Shots made by shots taken - all players in NBA season 2014-2015")
```

## How many have taken over 500 shots?
```{r}
player_shots_summary %>%
  filter(total_shots_taken > 500) %>%
  nrow
```

Who made the most shots?
```{r eval=FALSE}
ggplot(player_shots_summary, aes(x = reorder(PLAYER_NAME, total_shots_made), y = total_shots_made, fill = TEAM_NAME)) +
  geom_bar(stat = "identity") +
  guides(col = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Shots made by player")
``` 
Those stray lines represent players who play for multiple teams. We can ignore them for now, just interested in making sure the data looks reasonable.

OK, who were the top 50 players with the best shooting percentage of players who took more than 50 shots in the season?
```{r eval=FALSE}
top_50_shooting_percentage <- d %>%
  group_by(PLAYER_NAME) %>%
  summarize(total_shots_taken = n(),
            total_shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = total_shots_made / total_shots_taken,
            avg_shot_distance = mean(SHOT_DISTANCE)) %>%
  filter(total_shots_taken >= 50) %>%
  arrange(-shooting_percentage) %>%
  head(n = 50)

ggplot(top_50_shooting_percentage, aes(x = reorder(PLAYER_NAME, shooting_percentage), y = shooting_percentage)) +
    geom_bar(stat = "identity") +
    guides(col = FALSE) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle("Shots made by player")
```

Interesting, what kind of shots are these guys taking? Let's look a distance...
```{r eval=FALSE}
top50_names <- unique(top_50_shooting_percentage$PLAYER_NAME)

d %>%
  mutate(top_50 = ifelse(PLAYER_NAME %in% top50_names, TRUE, FALSE)) %>%
  group_by(PLAYER_NAME) %>%
  summarize(total_shots_taken = n(),
            total_shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = total_shots_made / total_shots_taken,
            avg_shot_distance = mean(SHOT_DISTANCE)) %>%
  filter(total_shots_taken >= 50) %>%
  ggplot(aes(x = reorder(PLAYER_NAME, shooting_percentage), y = shooting_percentage, fill = avg_shot_distance)) +
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle("Shots made by player")
```

# Section II - Populating streak info

```{r}
d_prev <- read.csv("d_prev_shots.csv")

if (!inGlobalEnv("d_prev")) {
  print("Writing d_prev_shots.csv... takes ~20min")
  ptm <- proc.time()
  ## Caution this takes ~5min to run
  d_prev <- populate_prev_shots(d)
  proc.time() - ptm  
  write.csv(d_prev, "d_prev_shots.csv")
}
```


## Indicator for streak of 3+
```{r}
d_prev <- d_prev %>%
  mutate(is_hot3 = curr_hit_streak >= 3,
         is_cold3 = curr_miss_streak >= 3)


d_prev %>%
  gather(streak_type, streak_value, c(curr_hit_streak, curr_miss_streak)) %>%
  group_by(streak_type, streak_value) %>%
  ggplot(aes(x=streak_value, y=SHOT_DISTANCE, col=streak_type)) +
  geom_point()

d_prev %>%
  gather(streak_type, streak_value, c(curr_hit_streak, curr_miss_streak)) %>%
  group_by(streak_type, streak_value) %>%
  summarize(shots_taken = n(),
            avg_shot_distance = mean(SHOT_DISTANCE),
            var_shot_distance = var(SHOT_DISTANCE),
            shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = shots_made / shots_taken,
            ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
            ci_upper = binom.test(shots_made, shots_taken)$conf.int[2]) %>%
  ggplot(aes(x=streak_value, y=shooting_percentage, col=streak_type)) +
  geom_point()
```


# Section III - Analysis

## In aggregate, how does the probability of a made shot change based on previously made shots?
```{r}
overall_cond_df <- d_prev %>%
  mutate(cond_probs_hit = ifelse(curr_hit_streak == 1, "1",
                                 ifelse(curr_hit_streak == 2, "2",
                                        ifelse(curr_hit_streak == 3, "3", "4+"))),
         cond_probs_miss = ifelse(curr_miss_streak == 1, "1",
                                 ifelse(curr_miss_streak == 2, "2",
                                        ifelse(curr_miss_streak == 3, "3", "4+")))) %>%
  gather(cond_probs_type, cond_probs_value, c(cond_probs_hit, cond_probs_miss)) %>%
  group_by(PLAYER_NAME, cond_probs_type, cond_probs_value) %>%
  summarize(shots_taken = n(),
            avg_shot_distance = mean(SHOT_DISTANCE),
            sd_shot_distance = sd(SHOT_DISTANCE),
            shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = shots_made / shots_taken,
            ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
            ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])

overall_row <- d_prev %>%
  mutate(cond_probs_type = "overall",
         cond_probs_value = 0) %>%
  group_by(cond_probs_type,cond_probs_value) %>%
  summarise(shots_taken=n(),
            avg_shot_distance = mean(SHOT_DISTANCE),
            sd_shot_distance = sd(SHOT_DISTANCE),
            shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = shots_made / shots_taken,
            ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
            ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])


overall_cond_df <- rbind(overall_cond_df, overall_row)
overall_cond_df$context <- c("made_1", "made_2", "made_3", "made_4+",
                             "missed_1", "missed_2", "missed_3", "missed_4+",
                             "overall")
row_order <- c("missed_4+", "missed_3", "missed_2", "missed_1", "overall", "made_1", "made_2", "made_3", "made_4+")
overall_cond_df <- overall_cond_df[match(row_order, overall_cond_df$context), ]
overall_cond_df$context <- factor(overall_cond_df$context, levels = c('made_4+',
                                                                      'made_3',
                                                                      'made_2',
                                                                      'made_1',
                                                                      'overall',
                                                                      'missed_1',
                                                                      'missed_2',
                                                                      'missed_3',
                                                                      'missed_4+'))

```


### Plot
```{r}
# K_Bryant <- d_prev %>%
#   filter(PLAYER_NAME == 'Kobe Bryant') %>%
#   mutate(prev_shot1 = curr_hit_streak >= 1,
#          prev_shot2 = curr_hit_streak >= 2,
#          prev_shot3 = curr_hit_streak >= 3) %>%
#   mutate(cond_probs_hit = ifelse(curr_hit_streak == 1, "1",
#                                  ifelse(curr_hit_streak == 2, "2",
#                                         ifelse(curr_hit_streak == 3, "3", "4+"))),
#          cond_probs_miss = ifelse(curr_miss_streak == 1, "1",
#                                  ifelse(curr_miss_streak == 2, "2",
#                                         ifelse(curr_miss_streak == 3, "3", "4+")))) %>%
#   gather(cond_probs_type, cond_probs_value, c(cond_probs_hit, cond_probs_miss)) %>%
#   group_by(cond_probs_type, cond_probs_value) %>%
#   summarize(shots_taken = n(),
#             avg_shot_distance = mean(SHOT_DISTANCE),
#             sd_shot_distance = sd(SHOT_DISTANCE),
#             shots_made = sum(SHOT_MADE_FLAG),
#             shooting_percentage = shots_made / shots_taken,
#             ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
#             ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
# 
#   
# bgtest <- bgtest(SHOT_MADE_FLAG ~ prev_shot3, order = 1, order.by = NULL, type = c("Chisq", "F"), data=K_Bryant)
# coeftest(bgtest)
# 
# dwtest(SHOT_MADE_FLAG ~ prev_shot, data = K_Bryant)

ggplot(overall_cond_df, aes(x = context, y = shooting_percentage)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Impact of previous shot on current shooting percentage\n
          All shots from NBA 2014-2015 season")
```


## Runs analysis

```{r}
players_over_100_shots <- d_prev %>%
  group_by(PLAYER_NAME) %>%
  summarise(num_shots = n()) %>%
  filter(num_shots >= 100) %>%
  select(PLAYER_NAME)
players_over_100_shots <- players_over_100_shots$PLAYER_NAME

run_test_df <- data.frame("PLAYER_NAME"=NA, "RUNS_STAT"=NA, "RUNS_PVALUE"=NA)
for (player in players_over_100_shots) {
  res <- calc_runs(subset(d_prev, as.character(PLAYER_NAME) == player)$SHOT_MADE_FLAG, pmethod="normal")
  statistic <- as.numeric(res[1])
  p.value <- as.numeric(res[2])
  run_test_df <- rbind(run_test_df, c("PLAYER_NAME"=player, "RUNS_STAT"=statistic, "RUNS_PVALUE"=p.value))
}
run_test_df <- run_test_df %>%
  filter(!is.na(PLAYER_NAME)) %>%
  mutate(RUNS_STAT = as.numeric(RUNS_STAT),
         RUNS_PVALUE = as.numeric(RUNS_PVALUE))
```

### plot
```{r}
run_test_df_plot <- run_test_df %>% 
  mutate(PLAYER_NAME = as.factor(PLAYER_NAME),
         p_value_col = ifelse(RUNS_PVALUE < 0.01, "green", "grey")) %>%
  arrange(-RUNS_STAT)

x_breaks <- levels(run_test_df_plot$PLAYER_NAME)[rep(c(TRUE,FALSE), length(run_test_df_plot$PLAYER_NAME)/2)]
ggplot(run_test_df_plot, aes(x=reorder(PLAYER_NAME, RUNS_STAT), y=RUNS_STAT, fill=p_value_col)) +
  geom_bar(stat="identity", position="identity") +
  scale_fill_manual(values=c("forest green", "grey")) +
  xlab("Players") +
  ylab("Z") +
  theme(axis.text.x = element_blank())


```


### Who had the longest runs (streaks)
```{r}
calc_runs(subset(d_prev, as.character(PLAYER_NAME) == "Al Jefferson")$SHOT_MADE_FLAG)
strruns(lawstat::runs.test(subset(d_prev, as.character(PLAYER_NAME) == "Al Jefferson")$SHOT_MADE_FLAG))



d_prev %>%
  filter(PLAYER_NAME %in% players_over_1000_shots) %>%
  group_by(PLAYER_NAME) %>%
  summarise(runs_test = runs.test(SHOT_MADE_FLAG))
  
d_prev %>%
  filter(curr_hit_streak > 10)

K_Bryant <- d_prev %>%
  filter(PLAYER_NAME == 'Kobe Bryant')
runs.test(K_Bryant$SHOT_MADE_FLAG)
```


```{r}
d_prev %>%
  gather(streak_type, streak_value, c(curr_hit_streak, curr_miss_streak)) %>%
ggplot(., aes(x=streak_value, y=SHOT_DISTANCE)) +
  geom_point(alpha=0.25) +
  geom_jitter(width=1) +
  facet_wrap(~streak_type) +
  geom_smooth(method="lm")
```

## lmer predict SHOT_DISTANCE
```{r}
library(lme4)
library(lmerTest)
summary(lmer(SHOT_DISTANCE ~ curr_hit_streak + curr_miss_streak + (1 | PLAYER_ID) + (1 | POSITION), data=d_prev))
```


```{r}
# d.raw_mini <- d[1:1000,]
# d.raw_mini2 <- populate_prev_shots(d.raw_mini)
# d.raw_mini2 %>% select(GAME_ID, PLAYER_NAME, PERIOD, TOTAL_TIME_REMAINING, time_between_shots, curr_hit_streak, curr_miss_streak) %>% View
# undebug(get_time_between_shots)
```


## Shooting percentage by shot number
```{r}
d_prev %>%
  group_by(shot_num) %>%
  filter(shot_num < 30) %>%
  summarize(shots_taken = n(),
            avg_shot_distance = mean(SHOT_DISTANCE),
            shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = shots_made / shots_taken,
            ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
            ci_upper = binom.test(shots_made, shots_taken)$conf.int[2]) %>%
  ggplot(aes(x = shot_num, y = shooting_percentage, fill = avg_shot_distance)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper))
```

Hmmm, this plot makes it look like accuracy stays steady, but `avg_shot_distance` is possibly incresing...

```{r}
d_prev %>%
  ggplot(aes(x = shot_num, y = SHOT_DISTANCE, col = as.factor(SHOT_MADE_FLAG))) +
    geom_point(alpha = 0.4)
```

Well, it actually isn't the case that people are taking longer shots later in the game - that was an artificat of increased variance with fewer sample at the higher number of shots.

## League-wide previous shots effects

### Data
```{r}
calc_prev_shot_effects <- function(df) {
  d_current <- df %>%
    summarize(shots_taken = n(),
              shots_made = sum(SHOT_MADE_FLAG),
              shooting_percentage = shots_made / shots_taken,
              ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
              ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_made_last_1 <- df %>%
      filter(made_last_1 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_made_last_2 <- df %>%
      filter(made_last_2 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_made_last_3 <- df %>%
      filter(made_last_3 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_made_last_4 <- df %>%
      filter(made_last_4 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_missed_last_1 <- df %>%
      filter(missed_last_1 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_missed_last_2 <- df %>%
      filter(missed_last_2 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_missed_last_3 <- df %>%
      filter(missed_last_3 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_missed_last_4 <- df %>%
      filter(missed_last_4 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  
  d_prev_shot_effects <- rbind(d_made_last_1,
                               d_made_last_2,
                               d_made_last_3,
                               d_made_last_4,
                               d_current,
                               d_missed_last_1,
                               d_missed_last_2,
                               d_missed_last_3,
                               d_missed_last_4) %>%
    mutate(context = c('made_1',
                       'made_2',
                       'made_3',
                       'made_4',
                       'overall',
                       'missed_1',
                       'missed_2',
                       'missed_3',
                       'missed_4'))
  
  d_prev_shot_effects$context <- factor(d_prev_shot_effects$context, levels = c('made_4',
                                                                                'made_3',
                                                                                'made_2',
                                                                                'made_1',
                                                                                'overall',
                                                                                'missed_1',
                                                                                'missed_2',
                                                                                'missed_3',
                                                                                'missed_4'))
  d_prev_shot_effects
}
```


```{r}

unique(d_prev$PLAYER_ID)

d_prev %>%
  group_by(PLAYER_ID) %>%
  summarize(shots = n())

d_prev %>% filter(PLAYER_ID == 203994) %>% calc_prev_shot_effects()


d_prev_shots_by_team <- plyr::ddply(d_prev, .variables = c("PLAYER_ID"), calc_prev_shot_effects)

ggplot(d_prev_shots_by_team, aes(x = context, y = shooting_percentage, fill = TEAM_ID)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~TEAM_ID)
  ggtitle("Impact of previous shot on current shooting percentage\n
          All shots from NBA 2014-2015 season\n
          By team")

```


### Plot
```{r}
d_prev_shot_effects <- calc_prev_shot_effects(d_prev)
ggplot(d_prev_shot_effects, aes(x = context, y = shooting_percentage)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Impact of previous shot on current shooting percentage\n
          All shots from NBA 2014-2015 season")
```

# Exploring previous shots
```{r}
library(lme4)
d_prev_small <- d_prev[1:10000,]
names(d_prev_small)


d_prev_small %>%
  mutate(hit_last_3_or_more = curr_hit_streak >= 3,
         missed_last_3_or_more = curr_miss_streak >= 3)
summary(lmer(SHOT_MADE_FLAG ~ curr_hit_streak + SHOT_DISTANCE * SHOT_TYPE, family="binomial", data=d_prev_small))

summary(glm(SHOT_MADE_FLAG ~ curr_hit_streak + SHOT_DISTANCE + ACTION_TYPE, family="binomial", data=d_prev_small))
```


# Modeling
```{r}
summary(glm(SHOT_MADE_FLAG ~ SHOT_TYPE +
              MINUTES_REMAINING +
              SHOT_DISTANCE + 
              made_last_1 +
              made_last_2 +
              made_last_3 + 
              made_last_4,
            family = "binomial", na.action = "na.omit", data = d_prev))
```

