---
title: "hotHandAnalysis"
author: "Ben Peloquin"
date: "May 6, 2016"
output: 
  html_document:
    toc: true
    toc_depth: 5
---

Prelims
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
rm(list = ls())
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
setwd('/Users/benpeloquin/Desktop/Spring2016/Stats267/hot_hand/analysis')
source("analysis_helpers.R")
data_path <- "../data/"
```

# Section I - General data orientation

Read in data. First check for consolidated `full_season_data_payer_info.csv`, otherwise create. (See `get_data.py` for data collection code)
```{r data_read1}
## Note this takes appr 2min to run
d.raw <- read.csv(paste0(data_path, "full_season_data_player_info.csv"))

if (!inGlobalEnv("d.raw")) {
  files <- list.files(path)
  d.raw <- data.frame()
  for (f in files) {
    d <- read.csv(paste0(path, f))
    d.raw <- rbind(d.raw, d)
  }
  
  d.player_info <- read.csv("player_info/players_info.csv") %>%
    select(player_number, age, weight, height, years_in_league, player_id, position)
  names(d.player_info) <- c("PLAYER_NUMBER", "AGE", "WEIGHT", "HEIGHT", "YEARS_IN_LEAGUE", "PLAYER_ID", "POSITION")
  
  d.raw <- full_join(d.raw, d.player_info, by = 'PLAYER_ID')
  write.csv(d.raw, "full_season_data_player_info.csv")
}
```

## Pre-processing

We have `r dim(d.raw)[1]` observations with `r dim(d.raw)[2]` predictor variables.

There are a number of predictors we might want to recode. Looks like `GAME_ID`, `GAME_EVENT_ID`, `PLAYER_ID`, `TEAM_ID`, `PERIOD`, `SHOT_ATTEMPTED_FLAG`, and `SHOT_MADE_FLAG` should all be coded as factors (currently coded as ints). Let's also convert `SHOT_TYPE` to an integer value...
```{r data_convert_types}
d <- d.raw %>%
  mutate(GAME_ID = as.factor(GAME_ID),
         GAME_EVENT_ID = as.factor(GAME_EVENT_ID),
         PLAYER_ID = as.factor(PLAYER_ID),
         TEAM_ID = as.factor(TEAM_ID),
         PERIOD = as.factor(PERIOD),
         SHOT_VALUE = ifelse(SHOT_TYPE == "2PT Field Goal", 2, 3),
         TOTAL_TIME_REMAINING = SECONDS_REMAINING + 60 * MINUTES_REMAINING)
```

Let's do some preliminary exploring just to make sure this data makes sense...

## Basic data orientation / summary

### Number of shots by player
```{r plot_shots_taken_shots_made}
player_shots_summary <- d %>%
  group_by(PLAYER_NAME, TEAM_NAME) %>%
  summarize(total_shots_taken = n(),
            total_shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = total_shots_made / total_shots_taken)

ggplot(player_shots_summary, aes(x = total_shots_taken, y = total_shots_made, col = TEAM_NAME)) +
  geom_point(alpha = 0.5, size = 4) +
  geom_smooth(method = "lm", aes(group=1)) + 
  guides(col = FALSE) +
  ggtitle("Shots made by shots taken - all players in NBA season 2014-2015")
```

### Total number of shots taken
```{r data_summary_num_shots_taken}
sum(player_shots_summary$total_shots_taken)
```

### Total number of shots made
```{r data_summary_num_shots_made}
sum(player_shots_summary$total_shots_made)
```

### Number of players
```{r data_summary_num_players}
length(unique(d$PLAYER_ID)) == length(unique(d$PLAYER_NAME))
length(unique(d$PLAYER_ID))
```

### Number of teams
```{r data_summary_num_teams}
length(unique(d$TEAM_ID))
```

Looks good, there are 30 teams

# Section II - Populating streak info

Check for pre-consolidated streak info in `d_prev_shots.csv`. If not write to new .csv. See (https://github.com/benpeloquin7/hot_hand/blob/master/analysis/analysis_helpers.R) for `populate_prev_shot()` helper functionality.
```{r data_streak_info}
d_prev <- read.csv(paste0(data_path, "d_prev_shots.csv"))

if (!inGlobalEnv("d_prev")) {
  print("Writing d_prev_shots.csv... takes ~20min")
  ptm <- proc.time()
  ## Caution this takes ~5min to run
  d_prev <- populate_prev_shots(d)
  proc.time() - ptm  
  write.csv(d_prev, "d_prev_shots.csv")
}
```

```{r eval=FALSE, echo=FALSE}
d_prev <- d_prev %>%
  mutate(is_hot3 = curr_hit_streak >= 3,
         is_cold3 = curr_miss_streak >= 3)


d_prev %>%
  gather(streak_type, streak_value, c(curr_hit_streak, curr_miss_streak)) %>%
  group_by(streak_type, streak_value) %>%
  ggplot(aes(x=streak_value, y=SHOT_DISTANCE, col=streak_type)) +
  geom_point()

d_prev %>%
  gather(streak_type, streak_value, c(curr_hit_streak, curr_miss_streak)) %>%
  group_by(streak_type, streak_value) %>%
  summarize(shots_taken = n(),
            avg_shot_distance = mean(SHOT_DISTANCE),
            var_shot_distance = var(SHOT_DISTANCE),
            shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = shots_made / shots_taken,
            ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
            ci_upper = binom.test(shots_made, shots_taken)$conf.int[2]) %>%
  ggplot(aes(x=streak_value, y=shooting_percentage, col=streak_type)) +
  geom_point()
```


# Section III - Analysis

## Question 1: Global conditional probabilities

### Data prep

This makes a test of the GVT's finding that an arbitrary players' shooting percentage should increase conditioned on a previously made shot.
```{r analysis1_cond_probs}
## Conditional probs
overall_cond_df <- d_prev %>%
  mutate(cond_probs_hit = ifelse(curr_hit_streak == 1, "1",
                                 ifelse(curr_hit_streak == 2, "2",
                                        ifelse(curr_hit_streak == 3, "3", "4+"))),
         cond_probs_miss = ifelse(curr_miss_streak == 1, "1",
                                 ifelse(curr_miss_streak == 2, "2",
                                        ifelse(curr_miss_streak == 3, "3", "4+")))) %>%
  gather(cond_probs_type, cond_probs_value, c(cond_probs_hit, cond_probs_miss)) %>%
  group_by(cond_probs_type, cond_probs_value) %>%
  summarize(shots_taken = n(),
            avg_shot_distance = mean(SHOT_DISTANCE),
            sd_shot_distance = sd(SHOT_DISTANCE),
            shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = shots_made / shots_taken,
            ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
            ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])

## Overall probs
overall_row <- d_prev %>%
  mutate(cond_probs_type = "overall",
         cond_probs_value = 0) %>%
  group_by(cond_probs_type, cond_probs_value) %>%
  summarise(shots_taken=n(),
            avg_shot_distance = mean(SHOT_DISTANCE),
            sd_shot_distance = sd(SHOT_DISTANCE),
            shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = shots_made / shots_taken,
            ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
            ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])

## Combine data
overall_cond_df <- rbind(overall_cond_df, overall_row)
overall_cond_df$context <- c("made_1", "made_2", "made_3", "made_4+",
                             "missed_1", "missed_2", "missed_3", "missed_4+",
                             "overall")
row_order <- c("missed_4+", "missed_3", "missed_2", "missed_1", "overall", "made_1", "made_2", "made_3", "made_4+")

## Format for plotting
overall_cond_df <- overall_cond_df[match(row_order, overall_cond_df$context), ]
overall_cond_df$context <- factor(overall_cond_df$context, levels = c('made_4+',
                                                                      'made_3',
                                                                      'made_2',
                                                                      'made_1',
                                                                      'overall',
                                                                      'missed_1',
                                                                      'missed_2',
                                                                      'missed_3',
                                                                      'missed_4+'))
## Populate standard errors
overall_cond_df <- overall_cond_df %>%
  mutate(se = ci_upper - ci_lower / 3.92)
```

### t-test

Crude, but makes a test of the hypothesis
```{r analysis1_cond_probs_ttest}
## t-test for made prev shot
## ------------------------
n1 <- overall_cond_df[4, ]$shots_taken       ## missed previous
n2 <- overall_cond_df[6, ]$shots_taken       ## made previous
sd1 <- overall_cond_df[4, ]$se / sqrt(n1)
sd2 <- overall_cond_df[6, ]$se / sqrt(n2)
shooting_per1 <- overall_cond_df[4, ]$shooting_percentage
shooting_per2 <- overall_cond_df[6, ]$shooting_percentage
se <- sqrt(sd1^2/n1 + sd2^2/n2)
t <- (shooting_per2 - shooting_per1) / se
p <- pt(t, n1 + n2 - 2)
cat(paste0("t = ", t), "\n", paste0("p < 0.001"))
```

### Plot
```{r analysis1_cond_probs_plot}
overall_cond_df %>%
  select(shooting_percentage, ci_lower, ci_upper)

ggplot(overall_cond_df, aes(x = context, y = shooting_percentage)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Impact of previous shot on current shooting percentage\n
          All shots from NBA 2014-2015 season")
```

## Question 2: Wald-Wolfowitz run test

Tests whether the observed number of runs differs from what would be expected by chance alone. Calculates (observed - expected) number of runs so significantly negative z-scores indicate significantly streaky players and significnatly positive z-scores indicate players who have fewer streaks than chance predictions.

See (https://github.com/benpeloquin7/hot_hand/blob/master/analysis/analysis_helpers.R) for Wald-Wolfowitz implementation.

### Data prep
```{r analysis2_WWruns_analsysis_data}
## Limit analysis to players with over 100 shots
players_over_100_shots <- d_prev %>%
  group_by(PLAYER_NAME) %>%
  summarise(num_shots = n()) %>%
  filter(num_shots >= 100) %>%
  select(PLAYER_NAME)
players_over_100_shots <- players_over_100_shots$PLAYER_NAME

## Num players
length(players_over_100_shots)
```

### WW runs for each player
```{r analysis2_WWruns_analsysis_test}
run_test_df <- data.frame("PLAYER_NAME"=NA, "RUNS_STAT"=NA, "RUNS_PVALUE"=NA)
for (player in players_over_100_shots) {
  res <- calc_runs(subset(d_prev, as.character(PLAYER_NAME) == player)$SHOT_MADE_FLAG, pmethod="normal")
  statistic <- as.numeric(res[1])
  p.value <- as.numeric(res[2])
  run_test_df <- rbind(run_test_df, c("PLAYER_NAME"=player, "RUNS_STAT"=statistic, "RUNS_PVALUE"=p.value))
}

run_test_df <- run_test_df %>%
  filter(!is.na(PLAYER_NAME)) %>%
  mutate(RUNS_STAT = as.numeric(RUNS_STAT),
         RUNS_PVALUE = as.numeric(RUNS_PVALUE))
```

### Plot
```{r analysis2_WWruns_analsysis_plot}
run_test_df_plot <- run_test_df %>% 
  mutate(PLAYER_NAME = as.factor(PLAYER_NAME),
         p_value_col = ifelse(RUNS_PVALUE < 0.01, "p<0.01", "non-sig")) %>%
  arrange(-RUNS_STAT)

x_breaks <- levels(run_test_df_plot$PLAYER_NAME)[rep(c(TRUE,FALSE), length(run_test_df_plot$PLAYER_NAME)/2)]
ggplot(run_test_df_plot, aes(x=reorder(PLAYER_NAME, RUNS_STAT), y=RUNS_STAT, fill=p_value_col)) +
  geom_bar(stat="identity", position="identity") +
  scale_fill_manual(values=c("grey", "forest green")) +
  xlab("Players") +
  ylab("Z") +
  theme(axis.text.x = element_blank()) +
  ggtitle(paste0("Wald-Wolfowitz run test", "\n",
          "All players from the 2014-15 season with 100+ shots"))
```

### Num players with p < 0.001
```{r}
run_test_df_plot %>%
  filter(RUNS_PVALUE < 0.01) %>%
  select(PLAYER_NAME, RUNS_STAT, RUNS_PVALUE) %>%
  knitr::kable(., digits=4)
```

## Question 3: Does prev shot predict shot distance?

M&S claim that ing-game is inappropriate for analysis because of confounds such as shot difficulty. We examine one aspect - does streak shooting (or missing) predict `SHOT_DISTANCE`?

### Data prep
```{r analysis3_shot_distance_data}
shot_dist_plot <- d_prev %>%
  gather(streak_type, streak_value, c(curr_hit_streak, curr_miss_streak)) %>%
  mutate(streak_type = ifelse(streak_type == "curr_hit_streak", streak_type, "_curr_miss_streak"))
```

### Plot
```{r analysis3_shot_distance_plot}
ggplot(shot_dist_plot, aes(x=streak_value, y=SHOT_DISTANCE, col=streak_type)) +
  # geom_point() +
  geom_jitter(alpha=0.25, size=0.6, width=1) +
  facet_wrap(~streak_type) +
  geom_smooth(method="lm") +
  ggtitle(paste0("Shot distance by hit/miss streaks", "\n",
          "All shots from the 2014-15 season"))
```

### Test

Fitting a mixed effects model regressing `SHOT_DISTANCE` on `curr_hit_streak`,  `curr_miss_streak` and `player` and `team` level random effects.
```{r analysis3_shot_distance_test}
# names(d_prev)
sum <- summary(lmer(SHOT_DISTANCE ~
               curr_hit_streak + curr_miss_streak + (1 | PLAYER_ID) + (1 | TEAM_NAME), data=d_prev))
knitr::kable(sum$coefficients, digits=3)
```

Looks like hot and cold streaks do predict `SHOT_DISTANCE`. This is the first evidence that there might be Hot Hand like effects. Of course this is only indirect. We will certainly want to control of `SHOT_DISTANCE` in our final analysis. 

## Question 4: Controlling in-game vars.

```{r analysis4_predict_shots_data}
predictors <- data.frame("SHOT" = c("curr_streak",
                                    "shot type",
                                    "shot distance",
                                    "shot value"),
                         "PLAYER" = c("height",
                                      "weight",
                                      "age",
                                      "years in league"),
                         "GAME" = c("total time remaining",
                                    "period",
                                    rep(NA, 2)))
knitr::kable(predictors)
```


```{r analysis4_predict_shots_test}
# names(d_prev)
sum <- summary(glm(SHOT_MADE_FLAG ~
                     curr_hit_streak + curr_miss_streak + SHOT_VALUE +
                     TOTAL_TIME_REMAINING + SHOT_DISTANCE + AGE + HEIGHT +
                     WEIGHT + ACTION_TYPE, family="binomial", data=d_prev))
knitr::kable(sum$coefficients[2:3, ], digits=3)
```


# Section IV: Other

### Shooting percentage by shot number
```{r eval=FALSE}
d_prev %>%
  group_by(shot_num) %>%
  filter(shot_num < 30) %>%
  summarize(shots_taken = n(),
            avg_shot_distance = mean(SHOT_DISTANCE),
            shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = shots_made / shots_taken,
            ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
            ci_upper = binom.test(shots_made, shots_taken)$conf.int[2]) %>%
  ggplot(aes(x = shot_num, y = shooting_percentage, fill = avg_shot_distance)) +
    geom_bar(stat = "identity") +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper))
```

Hmmm, this plot makes it look like accuracy stays steady, but `avg_shot_distance` is possibly incresing...

```{r eval=FALSE}
d_prev %>%
  ggplot(aes(x=shot_num, y=SHOT_DISTANCE, col=as.factor(SHOT_MADE_FLAG))) +
    geom_point(alpha=0.4) +
    geom_jitter(width=0.5) +
    geom_smooth(method="lm")
```

Well, it actually isn't the case that people are taking longer shots later in the game - that was an artificat of increased variance with fewer sample at the higher number of shots.

## League-wide previous shots effects

### Data
```{r eval=FALSE}
calc_prev_shot_effects <- function(df) {
  d_current <- df %>%
    summarize(shots_taken = n(),
              shots_made = sum(SHOT_MADE_FLAG),
              shooting_percentage = shots_made / shots_taken,
              ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
              ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_made_last_1 <- df %>%
      filter(made_last_1 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_made_last_2 <- df %>%
      filter(made_last_2 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_made_last_3 <- df %>%
      filter(made_last_3 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_made_last_4 <- df %>%
      filter(made_last_4 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_missed_last_1 <- df %>%
      filter(missed_last_1 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_missed_last_2 <- df %>%
      filter(missed_last_2 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_missed_last_3 <- df %>%
      filter(missed_last_3 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  d_missed_last_4 <- df %>%
      filter(missed_last_4 == TRUE) %>%
      summarize(shots_taken = n(),
                shots_made = sum(SHOT_MADE_FLAG),
                shooting_percentage = shots_made / shots_taken,
                ci_lower = binom.test(shots_made, shots_taken)$conf.int[1],
                ci_upper = binom.test(shots_made, shots_taken)$conf.int[2])
  
  d_prev_shot_effects <- rbind(d_made_last_1,
                               d_made_last_2,
                               d_made_last_3,
                               d_made_last_4,
                               d_current,
                               d_missed_last_1,
                               d_missed_last_2,
                               d_missed_last_3,
                               d_missed_last_4) %>%
    mutate(context = c('made_1',
                       'made_2',
                       'made_3',
                       'made_4',
                       'overall',
                       'missed_1',
                       'missed_2',
                       'missed_3',
                       'missed_4'))
  
  d_prev_shot_effects$context <- factor(d_prev_shot_effects$context, levels = c('made_4',
                                                                                'made_3',
                                                                                'made_2',
                                                                                'made_1',
                                                                                'overall',
                                                                                'missed_1',
                                                                                'missed_2',
                                                                                'missed_3',
                                                                                'missed_4'))
  d_prev_shot_effects
}
```


```{r eval=FALSE}

unique(d_prev$PLAYER_ID)

d_prev %>%
  group_by(PLAYER_ID) %>%
  summarize(shots = n())

d_prev %>% filter(PLAYER_ID == 203994) %>% calc_prev_shot_effects()


d_prev_shots_by_team <- plyr::ddply(d_prev, .variables = c("PLAYER_ID"), calc_prev_shot_effects)

ggplot(d_prev_shots_by_team, aes(x = context, y = shooting_percentage, fill = TEAM_ID)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~TEAM_ID)
  ggtitle("Impact of previous shot on current shooting percentage\n
          All shots from NBA 2014-2015 season\n
          By team")

```


### Plot
```{r eval=FALSE}
d_prev_shot_effects <- calc_prev_shot_effects(d_prev)
ggplot(d_prev_shot_effects, aes(x = context, y = shooting_percentage)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Impact of previous shot on current shooting percentage\n
          All shots from NBA 2014-2015 season")
```

# Exploring previous shots
```{r eval=FALSE}
d_prev_small <- d_prev[1:10000,]
names(d_prev_small)


d_prev_small %>%
  mutate(hit_last_3_or_more = curr_hit_streak >= 3,
         missed_last_3_or_more = curr_miss_streak >= 3)
summary(lmer(SHOT_MADE_FLAG ~ curr_hit_streak + SHOT_DISTANCE * SHOT_TYPE, family="binomial", data=d_prev_small))

summary(glm(SHOT_MADE_FLAG ~ curr_hit_streak + SHOT_DISTANCE + ACTION_TYPE, family="binomial", data=d_prev_small))
```

## Modeling
```{r eval=FALSE}
summary(glm(SHOT_MADE_FLAG ~ SHOT_TYPE +
              MINUTES_REMAINING +
              SHOT_DISTANCE + 
              made_last_1 +
              made_last_2 +
              made_last_3 + 
              made_last_4,
            family = "binomial", na.action = "na.omit", data = d_prev))
```

## How many have taken over 500 shots?
```{r eval=FALSE}
player_shots_summary %>%
  filter(total_shots_taken > 500) %>%
  nrow
```

Who made the most shots?
```{r eval=FALSE}
ggplot(player_shots_summary, aes(x = reorder(PLAYER_NAME, total_shots_made), y = total_shots_made, fill = TEAM_NAME)) +
  geom_bar(stat = "identity") +
  guides(col = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Shots made by player")
``` 

Those stray lines represent players who play for multiple teams. We can ignore them for now, just interested in making sure the data looks reasonable.

OK, who were the top 50 players with the best shooting percentage of players who took more than 50 shots in the season?
```{r eval=FALSE}
top_50_shooting_percentage <- d %>%
  group_by(PLAYER_NAME) %>%
  summarize(total_shots_taken = n(),
            total_shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = total_shots_made / total_shots_taken,
            avg_shot_distance = mean(SHOT_DISTANCE)) %>%
  filter(total_shots_taken >= 50) %>%
  arrange(-shooting_percentage) %>%
  head(n = 50)

ggplot(top_50_shooting_percentage, aes(x = reorder(PLAYER_NAME, shooting_percentage), y = shooting_percentage)) +
    geom_bar(stat = "identity") +
    guides(col = FALSE) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle("Shots made by player")
```

Interesting, what kind of shots are these guys taking? Let's look a distance...
```{r eval=FALSE}
top50_names <- unique(top_50_shooting_percentage$PLAYER_NAME)

d %>%
  mutate(top_50 = ifelse(PLAYER_NAME %in% top50_names, TRUE, FALSE)) %>%
  group_by(PLAYER_NAME) %>%
  summarize(total_shots_taken = n(),
            total_shots_made = sum(SHOT_MADE_FLAG),
            shooting_percentage = total_shots_made / total_shots_taken,
            avg_shot_distance = mean(SHOT_DISTANCE)) %>%
  filter(total_shots_taken >= 50) %>%
  ggplot(aes(x = reorder(PLAYER_NAME, shooting_percentage), y = shooting_percentage, fill = avg_shot_distance)) +
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle("Shots made by player")
```
